---
title: "p8105_hw1_sz3318"
author: "Shike Zhang"
date: "2024-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
```{r}
 data("penguins", package = "palmerpenguins")
# dataset size
n_rows <- nrow(penguins)
n_cols <- ncol(penguins)

# Calculate mean 
mean_flipper_length <- mean(penguins$flipper_length_mm, na.rm = TRUE)

```
## Describtion:
The penguins dataset contains 344 n_rows rows and 8 n_cols columns. Some important variables include:

 * species: the species of penguins (Adelie, Chinstrap, Gentoo).
 * bill_length_mm: the length of the penguin's bill in millimeters.
 * flipper_length_mm: the length of the penguin's flippers in millimeters.
 
The average flipper length is 200.915 mean_flipper_length mm.

## Scatterplot
```{r}
library(ggplot2)

# Create scatterplot
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  theme_minimal() +
  labs(
    title = "Flipper Length vs Bill Length",
    x = "Bill Length (mm)",
    y = "Flipper Length (mm)"
  )

ggsave("scatterplot_penguins.png", width = 6, height = 4)
```
# Problem 2

```{r}
# Set seed for reproducibility
set.seed(123)

# Create the data frame
df <- data.frame(
  normal_sample = rnorm(10),                              # a random sample of size 10 from a standard Normal distribution
  logical_vector = rnorm(10) > 0,                         # a logical vector indicating whether elements of the sample are greater than 0
  character_vector = as.character(1:10),                  # a character vector of length 10
  factor_vector = factor(sample(c("A", "B", "C"), 10, replace = TRUE))  # a factor vector of length 10, with 3 different factor “levels”
)
df

```

```{r}
# Calculate the means (what works and what doesn't)
mean(df$normal_sample)    # Works
mean(df$logical_vector)   # Works (TRUE = 1, FALSE = 0)
```
## Describtion:
The mean of the normal_sample is 0.0746 and logical_vector is 0.7. The logical_vector can be interpreted as 1 for TRUE and 0 for FALSE. However, the character_vector and factor_vector cannot be directly averaged because they are not numeric.

```{r}
# Coercion to numeric
as.numeric(df$logical_vector)
as.numeric(df$character_vector)
as.numeric(df$factor_vector)
```

Logical Vector (as.numeric(df$logical_vector)):
The output [1 1 1 1 0 1 1 0 1 0] shows that TRUE is converted to 1 and FALSE is converted to 0. This helps explain why calculating the mean of a logical vector works

Character Vector (as.numeric(df$character_vector)):
The output [1 2 3 4 5 6 7 8 9 10] shows that the character vector was converted to its numeric representation. This works because the original character vector consists of strings that represent numbers "1", "2", "3". 

Factor Vector (as.numeric(df$factor_vector)):
The output [3 2 1 3 1 1 2 3 3 1] indicates that R converts factor levels to their internal numeric codes. This is helpful because we can averaging the underlying codes of the factor levels, not the categorical labels themselves.
